---
title: "Enhancing Customer Loyalty and Coupon Redemption Strategy for Waitrose"
output: word_document
date: "2023-07-15"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Enhancing Customer Loyalty and Coupon Redemption Strategy for Waitrose.


# ----------------------------------------------------------------------------------------------------
# ----------------------------------------------------------------------------------------------------

# Load the necessary packages
```{r}
library(readxl) # For importing data from excel
library(tidyverse) # For performing data manipulation
library(ggplot2) # For visualizing the dataset
library(lubridate) # For working with dates
library(car) # For modelling
library(stats) # For checking regression assumptions i.e the qqplot
```


# ----------------------------------------------------------------------------------------------------
# ----------------------------------------------------------------------------------------------------

# Importing dataset
```{r}
campaign_descriptions <- read.csv("C:/Users/User/Desktop/Dataset/campaign_descriptions.csv")
campaigns <- read.csv("C:/Users/User/Desktop/Dataset/campaigns.csv")
coupon_redemptions <- read.csv("C:/Users/User/Desktop/Dataset/coupon_redemptions.csv")
coupons <- read.csv("C:/Users/User/Desktop/Dataset/coupons.csv")
demographics <- read.csv("C:/Users/User/Desktop/Dataset/demographics.csv")
products <- read.csv("C:/Users/User/Desktop/Dataset/products.csv")
promotions <- read.csv("C:/Users/User/Desktop/Dataset/promotions.csv")
transactions <- read.csv("C:/Users/User/Desktop/Dataset/transactions.csv")
```


# exploring the first few observations in each data file
```{r}
head(campaign_descriptions)
head(campaigns)
head(coupon_redemptions)
head(coupons)
head(demographics)
head(products)
head(promotions)
head(transactions)
```
# --------------------------------------------------------------------------------------------------

# Data Exploration
```{r}
# Checking the structure of the campain description data
str(campaign_descriptions)
```

```{r}
str(campaigns)
```

```{r}
str(coupon_redemptions)
```

```{r}
str(coupons)
```

```{r}
str(demographics)
```

```{r}
str(products)
```

```{r}
str(promotions)
```

```{r}
str(transactions)
```

# --------------------------------------------------------------------------------------------------
# Data Cleaning
# Checking for missing values each data file
```{r}
colSums(is.na(head(campaign_descriptions)))
```

```{r}
colSums(is.na(head(campaigns)))
```

```{r}
colSums(is.na(head(coupon_redemptions)))
```

```{r}
colSums(is.na(head(coupons)))
```

```{r}
colSums(is.na(head(demographics)))
```

```{r}
colSums(is.na(head(products)))
```

```{r}
products2<-na.omit(products)
colSums(is.na(head(products2)))
```

```{r}
colSums(is.na(head(campaigns)))
```

```{r}
colSums(is.na(head(promotions)))
```

```{r}
colSums(is.na(head(transactions)))
```



# ---------------------------------------------------------------------------------------------------
# ---------------------------------------------------------------------------------------------------

# Exploratory data analysis

```{r}
summary(campaign_descriptions)

```

```{r}
summary(campaigns)
```

```{r}
summary(coupon_redemptions)
```

```{r}
summary(coupons)
```

```{r}
summary(demographics)
```

```{r}
summary(products)
```

```{r}
summary(promotions)
```

```{r}
summary(transactions)
```


# ----------------------------------------------------------------------------------------------------
# --------------------------------------------------------------------------------------------------

## Exploratiry data analysis and Visualizations 
## How many customers are spending more over time? Less overtime? Describe these customers.
```{r}
# Calculate total sales for each customer
customer_sales <- transactions %>%
  group_by(household_id) %>%
  summarise(total_sales = sum(sales_value))
head(customer_sales) 
```

```{r}
# Line plot for the total_sales more over rank
ggplot(customer_sales, aes(x =household_id , y = total_sales)) +
  geom_line(stat = "identity", fill = "blue") +
  labs(title = "total_sales more over household_id", x = "household_id", y = "total_sales")

```

```{r}
# Identify customers spending more over time
increasing_customers <- customer_sales %>%
  arrange(total_sales) %>%
  mutate(rank = row_number())
head(increasing_customers)
```

```{r}
# Line plot for the total_sales more over rank
ggplot(increasing_customers, aes(x =rank , y = total_sales)) +
  geom_line(stat = "identity", fill = "steelblue") +
  labs(title = "total_sales more over rank", x = "rank", y = "total_sales")

```


```{r}
# Describe customers spending more over time
top_customers <- increasing_customers %>%
  filter(rank <= 10)  # Change the number as per your preference
head(top_customers)
```

```{r}
# Line graph plot for the total_sales more over rank
ggplot(top_customers, aes(x =rank , y = total_sales)) +
  geom_line(stat = "identity", fill = "steelblue") +
  labs(title = "customers spending more over time", x = "rank", y = "Customer spending")
```

```{r}
# Describe customers spending less over time
bottom_customers <- decreasing_customers %>%
  filter(rank <= 10)  # Change the number as per your preference
head(bottom_customers)
```

```{r}
# Line plot for the customers spending less over time
ggplot(bottom_customers, aes(x =rank , y = total_sales)) +
  geom_line(stat = "identity", fill = "steelblue") +
  labs(title = "customers spending less over time", x = "rank", y = "Customer spending")
```

```{r}
# Which demographic factors (e.g. household size, presence of children, income) appear to affect customer spend? Engagement with certain categories?
ggplot(demographics, aes(x = marital_status, y = "")) +
  geom_bar(stat = "identity", fill = "green") +
  labs(title = "Customers marital status", x = "marital status", y = "Counts")
```


```{r}
# Bar plot for the Campaign type with the highest counts
ggplot(campaign_descriptions, aes(x = campaign_type, y = "")) +
  geom_bar(stat = "identity", fill = "steelblue") +
  labs(title = "Campaign type with the highest counts", x = "campaign_type", y = "Counts")


```

```{r}
# Plotting household_id over campaign_id
ggplot(campaigns, aes(x = campaign_id, y = household_id)) +
  geom_point() +
  labs(x = "campaign_id", y = "household_id", title = "household_id Over campaign_id")

```

```{r}

# Plotting coupon_upc over time
ggplot(coupon_redemptions, aes(x = redemption_date)) +
  geom_point(aes(y = coupon_upc, color = "coupon_upc")) +
  labs(x = "redemption_date", y = "Value", title = "coupon_upc Over redemption_date") +
  scale_color_manual(values = c("coupon_upc" = "blue"))

```

```{r}
# Plotting coupon_upc and household_id together
ggplot(coupon_redemptions, aes(x = redemption_date)) +
geom_point(aes(y = household_id, color = "household_id")) +
  labs(x = "redemption_date", y = "Value", title = "coupon_upc and household_id Over Time") +
  scale_color_manual(values = c("coupon_upc" = "blue", "household_id" = "red"))
```


```{r}
# Scatter plot of sales_value by quantity
ggplot(transactions, aes(x = transactions$sales_value, y = transactions$quantity)) +
  geom_point(stat = "identity") +
  labs(x = "sales_value", y = "quantity", title = "Quantity by sales value")


```

```{r}
# Boxplot of product_id by brand
ggplot(products, aes(x = brand, y = product_id)) +
  geom_boxplot() +
  labs(x = "brand", y = "product_id", title = "Product_id Distribution by brand")

```


```{r}
model1<-lm(sales_value~quantity+retail_disc+coupon_disc+coupon_match_disc, data = transactions)
summary(model1)
```















